Beta Diversity through Time and Space

### 2020-07-13 analyses/beta_diversity.Rmd
update repo

### Climate data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year),
         yday = yday(date)) %>%
  dplyr::select(site, year, date, yday, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")

gdd_total <- gdd %>%
  group_by(site, year) %>%
  summarize(gdd.total = max(gdd.cum))
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>% # turn elav.class into an ordered factor
  semi_join(read_csv("../processed_data/network.csv"), by = "site")

# %>%
#   left_join(gdd_total, by = "site")

### Plot total GDD
ggplot(site_data, aes(elev.mean, gdd.total, color = year)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  ylab("Total GDD")
```

### Load network and survey data and generate site.year and site.month.year webs
```{r}
# bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
#   dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)
# 
# fl_traits <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
#   dplyr::select(plant.sp, k.type) %>%
#   mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
#          k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))

# fl_tax <- read_csv("../processed_data/floral_tax.csv") 
# 
# breaks <- read_csv("./output/breaks.csv") %>%
#   mutate(year = factor(year))

net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.date, site, date, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  #full_join(bb_traits, by = c("bb.sp" = "bb.sp")) %>%
  left_join(site_data, by = "site") %>%
  #left_join(gdd, by = c("site", "date")) %>%
  #left_join(breaks, by = c("site", "year")) %>%
  #left_join(fl_traits) %>%
  #left_join(fl_tax, by = "plant.genus") %>%
  ungroup() %>%
  #mutate(phase = case_when(
    # gdd.cum < break1.gdd ~ "founding",
    # gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    # gdd.cum > break2.gdd ~ "reproductive"
  #)) %>%
  #mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  arrange(site, date)

@@

@@

beta_site <- betalinkr_multi(array_site, partition.st = TRUE, partition.rr = TRUE) %>%
  as_tibble() %>%
  rename(site1 = i) %>%
  rename(site2 = j) %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, site2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2)) %>%
  #left_join(site_pairwise_distance) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

write_csv(beta_site, "../analyses/output/beta_site.csv")

beta_site.year <- betalinkr_multi(array_site.year, partition.st = TRUE, partition.rr = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "year1"), sep = "_") %>%
  separate(j, into = c("site2", "year2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, year1, site2, year2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2))
```

# Beta diversity by elevation difference
```{r}
# No apparent differences across years, so we will use the pooled data
ggplot(filter(beta_site.year, year1 == year2), 
       aes(metric, value, color = year1)) +
  geom_boxplot() +
  facet_wrap(~metric, scales = "free")

# # And not much of a pattern in response to geographical distance
# ggplot(beta_site, 
#        aes(as.numeric(distance), value, color = metric)) +
#   geom_smooth()

# Interaction beta diversity by elevation difference
ggplot(beta_site, 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)
```

### Let's look at floral survey diversity patterns
```{r}
survey_bioData <- survey %>%
  filter(flower.cover > 0) %>%
  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
  group_by(site, plant.sp) %>%
  summarize(cover = sum(flower.cover)) %>%
  spread(plant.sp, cover) %>%
  replace(is.na(.), 0)

>>>>

### Climate data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year),
         yday = yday(date)) %>%
  dplyr::select(site, year, date, yday, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>% # turn elav.class into an ordered factor
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
```

### Load network and survey data and generate site.year and site.month.year webs
```{r}
net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.date, site, date, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  group_by(site, date) %>%
  #mutate(network.size = n()) %>%
  arrange(site, date)

@@

site.year_transects <- net %>%
  group_by(site, year) %>%
  summarize(transects = length(unique(date)))

site_transects <- net %>%
  group_by(site) %>%
  summarize(transects = length(unique(date)))

@@

beta_site <- betalinkr_multi(array_site, partition.st = TRUE, partition.rr = TRUE) %>%
  as_tibble() %>%
  rename(site1 = i) %>%
  rename(site2 = j)%>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  left_join(site_transects, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
  left_join(site_transects, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         transects.diff = abs(transects1 - transects2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, site2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, 
                           lat1, lon2, lat2, transects1, transects2, transects.diff)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

#write_csv(beta_site, "../analyses/output/beta_site.csv")

beta_site.year <- betalinkr_multi(array_site.year, partition.st = TRUE, partition.rr = TRUE, binary = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "year1"), sep = "_") %>%
  separate(j, into = c("site2", "year2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  left_join(site.year_transects, by = c("site1" = "site", "year1" = "year")) %>%
  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
  left_join(site.year_transects, by = c("site2" = "site", "year2" = "year")) %>%
  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         transects.diff = abs(transects1 - transects2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, year1, site2, year2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2,
                           transects1, transects2, transects.diff)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))
```

# Beta diversity by elevation difference
```{r}
# Interaction beta diversity by elevation difference (by year)
## This shows that the patterns are extremely consistent across years, which is not surprising, since they are driven overwhelmingly by floral turnover, and plants tend to be in the same place from year to year.
ggplot(filter(beta_site.year, year1 == year2 & 
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  facet_grid(metric.class ~ year1)

ggplot(filter(beta_site.year, year1 == year2 & 
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(transects.diff, value, color = metric)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  facet_grid(metric.class ~ year1, scales = "free_x")

# Interaction beta diversity by elevation difference, this time pooling across years
ggplot(filter(beta_site, metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

# How important is the difference in transects? Not as important as the difference in elevation.
ggplot(filter(beta_site, metric != "S" & 
                metric.class != "OS.partition"), 
       aes(transects.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

ggplot(beta_site, 
       aes(elev.diff, transects.diff)) +
  geom_point(alpha = 0.05) +
  geom_smooth()

### Okay, we'll just do partial mantel tests to partial out the effect of transects.diff; and we'll only do that for the aggregate measures because it seems a bit silly to doi it for the partitions  
WN_diff <- beta_site %>%
  filter(metric == "WN") %>%
  dplyr::select(site1, site2, value) %>%
  spread(site2, value) %>%
  dist()

WN_diff <- beta_site %>%
  filter(metric == "WN") %>%
  dplyr::select(site1, site2, value) %>%
  spread(site2, value) %>%
  dist()

ST_diff <- beta_site %>%
  filter(metric == "ST") %>%
  dplyr::select(site1, site2, value) %>%
  spread(site2, value) %>%
  dist()

OS_diff <- beta_site %>%
  filter(metric == "OS") %>%
  dplyr::select(site1, site2, value) %>%
  spread(site2, value) %>%
  dist()

elev_diff <- beta_site %>%
  dplyr::select(site1, site2, elev.diff) %>%
  unique() %>%
  spread(site2, elev.diff) %>%
  dist()

transects_diff <- beta_site %>%
  dplyr::select(site1, site2, transects.diff) %>%
  unique() %>%
  spread(site2, transects.diff) %>%
  dist()

WN_mantel <- mantel.partial(elev_diff, WN_diff, transects_diff)
ST_mantel <- mantel.partial(elev_diff, ST_diff, transects_diff)
OS_mantel <- mantel.partial(elev_diff, OS_diff, transects_diff)
```

### Let's look at floral survey diversity patterns
```{r}
survey_transects <- survey %>%
  group_by(site) %>%
  summarize(transects = length(unique(date)))

survey_bioData <- survey %>%
  filter(flower.cover > 0) %>%
  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
  group_by(site, plant.sp) %>%
  summarize(cover = sum(flower.cover)) %>%
  spread(plant.sp, cover) %>%
  replace(is.na(.), 0)

predData <- site_data %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon) %>%
  unique()
%>%
  left_join(survey_transects)
